# Results

```{r}
#| label: results-load
#| include: false

file_names <- 
  list.files(
    path = "figs", 
    pattern = "rdata", 
    full.names = TRUE,
    recursive = TRUE
  )

lapply(file_names, load, .GlobalEnv)

```

## Theoretical Model

In this section I summarise the reults from the interviews and literature assessment, and present the resulting theoretical model of NVS.

The theoretical model is presented as a diagram visualization, accompained with text explining the causal processes represented in the diagram.

## Sociological Analysis

In this section I show the identified social actors and how they relate with each other in the dispute over land, determined by NVS. The results will lso be represented as diagrams, that will show the relations between the social actors, and the capital of each.

## Statistical Analysis

In this section I present the results from the quantitative analysis, from the exploratory data analysis to the statistical model predictions and interpretation.

### Exploratory Data Analysis

In this section I analyse each variable of my model. 
It is important to note that all quantitative analysis were performed in the data stored in the base grid.
So the figures and tables do not show distributions of the raw observations/estimates, but the values that were transformed to fit the grid cells (@fig-base-grid). 

```{r}
#| label: fig-base-grid
#| fig-cap: Model grid, in which variables were transformed into.
#| fig-asp: 1
#| echo: false

viz_grid_area

```

All visualizations about the variables follows a common color scale, to facilitate their interpretation along the document (@fig-color-scale). The minimum values are represented by light grey color, values close to the median shows a yellow color, and maximum values as dark red color.

```{r}
#| label: fig-color-scale
#| fig-cap: Color scale used to visualize the variables properties.
#| fig-asp: 0.2
#| echo: false

color_scale

```

### Natural Vegetation Suppression

::: {.callout-important collapse="true"}

READ:

- https://ssrn.com/abstract=4063325
- https://doi.org/10.1086/597794 
- https://doi.org/10.1111/j.1475-682X.2007.00200.x

:::

```{r}
#| label: tbl-deforestation_summary
#| tbl-cap: "Summary table of natural vegetation suppression."
#| echo: false

htmltools::includeHTML("./figs/eda/deforestation_table.html")

```

```{r}
#| label: fig-deforestation-hist
#| fig-cap: "Distribution of deforestation in the Amazon and Cerrado. The horizontal scale is transformed to base 10 log."
#| fig-asp: 0.70
#| echo: false

deforestation_histogram

```

```{r}
#| label: fig-deforestation-cumsum
#| fig-cap: "Cumulative percent of the deforestation area."
#| fig-asp: 0.70
#| echo: false

deforestation_cumulative

```

```{r}
#| label: fig-deforestation-spatial-distribution
#| fig-cap: "Spatial distribution of the deforestation area."
#| fig-asp: 1
#| echo: false

deforestation_spatial

```

```{r}
#| label: fig-deforestation-timeseries
#| fig-cap: "Time series of deforestation area."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_timeseries

```

```{r}
#| label: fig-deforestation-colsum-natural
#| fig-cap: "Deforested area of different land cover."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_natural

```

```{r}
#| label: fig-deforestation-colsum-human
#| fig-cap: "Area of established LULC after deforestation."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_human

```

## Final results

The final results section will present the divergences and convergences between the qualitative and quantitative results, summarising the findings in tables and diagrams.

# References {.unnumbered}