# Results

```{r}
#| label: results-load
#| include: false

file_names <- 
  list.files(
    path = "figs", 
    pattern = "rdata", 
    full.names = TRUE,
    recursive = TRUE
  )

lapply(file_names, load, .GlobalEnv)

```

The study will present three main quantitative results based on quantitative data: predictions on vegetation suppression, land use transitions, and social and environmental variables. 
Our predictions on vegetation suppression will show the likelihood and extent of removal events, serving as a basis for predicting the land use changes of the affected grid cells. 
Finally, we will predict the impact of these findings on the environment and human well-being.

As part of the study, I will not only present the predictions generated by the models but also perform an in-depth analysis of the model's behavior. 
This analysis will provide a deeper understanding of the underlying mechanisms that drive the predictions and help to assess their reliability and accuracy.
Furthermore, based on my confidence in the models and their results, I will make inferences about the studied phenomena. 
This will allow me to go beyond the mere presentation of the predictions and offer insights into the relationships between the variables, the causes of the observed patterns, and the implications of the results for the environment and human well-being. 
However, it is important to note that I will only take this step if I feel confident in the models and their results, as making inferences requires a strong foundation of knowledge and understanding.

## Modelling variables

In this section I will analyse each variable of my model. 
It is important to note that all quantitative analysis were performed in the data stored in the base grid.
So the figures and tables do not show distributions of the raw observations/estimates, but the values that were transformed to fit the grid cells (@fig-base-grid). 

```{r}
#| label: fig-base-grid
#| fig-cap: Model grid, in which variables were transformed into.
#| fig-asp: 1
#| echo: false

viz_grid_area

```

All visualizations about the variables will follow a common color scale, to facilitate their interpretation along the document (@fig-color-scale). The minimum values will be represented by light grey color, values close to the median will show a yellow color, and maximum values as dark red colors.

```{r}
#| label: fig-color-scale
#| fig-cap: Color scale used to visualize the variables properties.
#| fig-asp: 0.2
#| echo: false

color_scale

```

### Deforestation

What is deforestation? Even worse, what is a forest? A recent research reviewed 65 reports, and found 15 definitions of deforestation, and 9 definitions of forests [@oca-2021].

::: {.callout-important collapse="true"}

TODO: 

- Choose between therms: deforestation or native vegetation suppression
- Tell in general the definitions for forests and deforestation (check the one proposed by oca-2021)
- Discuss about the problematic of forests definitions, which can have serious negative impacts to savannas, such as a good part of the Cerrado!
- Explain the definitions used by PRODES (INPE) to identify deforestation in the Cerrado and Amazon

READ:

- https://ssrn.com/abstract=4063325
- https://doi.org/10.1086/597794 
- https://doi.org/10.1111/j.1475-682X.2007.00200.x

:::

```{r}
#| label: tbl-deforestation_summary
#| tbl-cap: "Summary table of deforestation."
#| echo: false

htmltools::includeHTML("./figs/eda/deforestation_table.html")

```

```{r}
#| label: fig-deforestation-hist
#| fig-cap: "Distribution of deforestation in the Amazon and Cerrado. The horizontal scale is transformed to base 10 log."
#| fig-asp: 0.70
#| echo: false

deforestation_histogram

```

```{r}
#| label: fig-deforestation-cumsum
#| fig-cap: "Cumulative percent of the deforestation area."
#| fig-asp: 0.70
#| echo: false

deforestation_cumulative

```

```{r}
#| label: fig-deforestation-spatial-distribution
#| fig-cap: "Spatial distribution of the deforestation area."
#| fig-asp: 1
#| echo: false

deforestation_spatial

```

```{r}
#| label: fig-deforestation-timeseries
#| fig-cap: "Time series of deforestation area."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_timeseries

```

```{r}
#| label: fig-deforestation-colsum-natural
#| fig-cap: "Deforested area of different land cover."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_natural

```

```{r}
#| label: fig-deforestation-colsum-human
#| fig-cap: "Area of established LULC after deforestation."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_human

```

::: {.callout-important collapse="true"}

TODO: 

- Fix deforestation area
- Set the name of LULC classes
- Translate biomes names

:::

# References {.unnumbered}