---
css: styles.css
---

# Results

```{r}
#| label: results-load
#| include: false

file_names <- 
  list.files(
    path = "figs", 
    pattern = "rdata", 
    full.names = TRUE,
    recursive = TRUE
  )

lapply(file_names, load, .GlobalEnv)
```

## Theoretical Model

In this section I summarize the results from the interviews and literature assessment, and present the resulting theoretical model of NVS.

The theoretical model is presented as a diagram visualization, accompanied with text explaining the causal processes represented in the diagram.

## Sociological Analysis

In this section I show the identified social actors and how they relate with each other in the dispute over land, determined by NVS. The results will also be represented as diagrams, that will show the relations between the social actors, and the capital of each.

## Statistical Analysis

In this section I present the results from the quantitative analysis, from the exploratory data analysis to the statistical model predictions and interpretation.

### Exploratory Data Analysis

In this section I analyze each variable of my model. It is important to note that all quantitative analysis were performed in the data stored in the base grid. So the figures and tables do not show distributions of the raw observations/estimates, but the values that were transformed to fit the grid cells (@fig-base-grid).

```{r}
#| label: fig-base-grid
#| fig-cap: Model grid, in which variables were transformed into.
#| fig-asp: 1
#| echo: false
#| dev: "ragg_png"

viz_grid_area
```

All visualizations about the variables follows a common color scale, to facilitate their interpretation along the document (@fig-color-scale). The minimum values are represented by light gray color, values close to the median shows a yellow color, and maximum values as dark red color.

```{r}
#| label: fig-color-scale
#| fig-cap: Color scale used to visualize the variables properties.
#| fig-asp: 0.2
#| echo: false
#| dev: "ragg_png"

color_scale
```

#### Natural Vegetation Suppression

The NVS is the phenomena of interest of this project.

::: {.callout-important collapse="true"}
READ:

-   https://ssrn.com/abstract=4063325
-   https://doi.org/10.1086/597794
-   https://doi.org/10.1111/j.1475-682X.2007.00200.x
:::

In the grid of analysis, the minimum area of NVS is 2 ha for the Amazon and Cerrado, and the maximum area is 18000 ha a year for the Amazon, and 12000 ha a year in the Cerrado @tbl-deforestation-summary.

```{r}
#| label: tbl-deforestation-summary
#| tbl-cap: "Summary table of yearly natural vegetation suppression."
#| echo: false
#| dev: "ragg_png"

htmltools::includeHTML("./figs/eda/deforestation_table.html")
```

The distribution of NVS area in the grid cells is heavily skewed towards smaller areas, specially in the Amazon, which also shows a high frequency of a diverse range of values @fig-deforestation-hist. The distribution of NVS for the Cerrado is less skewed, and show more frequent values close to 1000 ha a year @fig-deforestation-hist.

```{r}
#| label: fig-deforestation-hist
#| fig-cap: "Distribution of natural vegetation suppression in the Amazon and Cerrado. The horizontal scale is transformed to base 10 log."
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

deforestation_histogram
```

The cumulative NVS area shows that grid cells with high values of NVS accounts for the majority of the total NVS in a year, for the Amazon NVS from 3000ha to 18000 ha, which are the 25% highest values, accounts for 79% of all NVS @fig-deforestation-cumsum. For the Cerrado, NVS from 900 ha to 12000 ha, the 25% highest values, accounts for 57% of the total NVS @fig-deforestation-cumsum.

```{r}
#| label: fig-deforestation-cumsum
#| fig-cap: "Cumulative percent of the natural vegetation suppression area."
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

deforestation_cumulative
```

The spatial distribution of NVS shows a concentration in some regions of the Amazon, specially in areas in proximity to the Cerrado biome [@fig-deforestation-spatial-distribution]. However, the distribution of NVS in the Cerrado shows a different pattern, where the NVS occurrence is spread in the whole biome [@fig-deforestation-spatial-distribution].

```{r}
#| label: fig-deforestation-spatial-distribution
#| fig-cap: "Spatial distribution of the natural vegetation suppression area."
#| fig-asp: 1
#| echo: false
#| dev: "ragg_png"

deforestation_spatial
```

The NVS time series shows periods of increase of NVS in the Amazon from 1985 to 2003, followed by a consistent drop of NVS area after 2003 [@fig-deforestation-timeseries]. For the Cerrado, the NVS area have been decreasing since the start of the time series, with some periods of increase in 2004 and 2013.

```{r}
#| label: fig-deforestation-timeseries
#| fig-cap: "Time series of natural vegetation suppression area."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

deforestation_timeseries
```

The natural vegetation types that suffered suppression are Forests, Grasslands, Savannas and Wetlands. In the Amazon, the majority of NVS occurs over Forests, and in the Cerrado it occurrence is predominant over Savannas [@fig-deforestation-colsum-natural].

```{r}
#| label: fig-deforestation-colsum-natural
#| fig-cap: "Natural vegetation suppression area of different land cover."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

deforestation_colsum_natural
```

The antropic land uses that followed after NVS are predominantly Pastures for both biomes [@fig-deforestation-colsum-human]. However, Mosaic of Uses and Temporary Crops share a significant amount of NVS area in the Cerrado [@fig-deforestation-colsum-human].

```{r}
#| label: fig-deforestation-colsum-human
#| fig-cap: "Area of established land use after natural vegetation suppression."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

deforestation_colsum_human
```

#### Conservation Units

The creation of conservation units is a main strategy to protect natural landscapes in Brazil. The creation of protected areas in Brazil dates from 1896, with the creation of the first National Park in São Paulo [@drummond-2009], however, it was only in 2000 that a law was approved to create a system to organize and regulate the UCs, named as the National System of Conservation Units (SNUC) [@drummond-2009; @borges-2009; @moreira-2021]. The SNUC comprises federal, state and municipal UCs of different characteristics [@borges-2009].

In the grid of analysis, the maximum area occupied by UC in a single cell is 78000 ha, which is the total cell area @fig-uc-summary. The majority of cells do not present any area of UC @fig-uc-summary.

```{r}
#| label: fig-uc-summary
#| fig-cap: "Summary table of Conservation Units."
#| fig-alt: ""
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

htmltools::includeHTML("./figs/eda/uc_table.html")
```

When analyzing the grid cells with any cover of UC, we observe that in the Amazon, most those grid cells are fully occupied by UC @fig-uc-hist. The situation is very different for the Cerrado, where most pixels occupied by UC shows small areas, indicating that small UC are more frequent for this biome @fig-uc-hist.

```{r}
#| label: fig-uc-hist
#| fig-cap: "Histogram of Conservation Units."
#| fig-alt: ""
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

uc_histogram
```

The spatial distributions of UC in the Amazon and Cerrado are very different. In the Amazon it is possible to observe a large number of big UC, while in Cerrado most of UC are smaller as scattered around the biome @fig-uc-spatial-distribution.

```{r}
#| label: fig-uc-spatial-distribution
#| fig-cap: "Spatial distribution of Conservation Units."
#| fig-alt: ""
#| fig-asp: 1
#| echo: false
#| dev: "ragg_png"

uc_spatial
```

The time series of accumulated UC area over time shows a steady increase from 1985 to 2010, followed by a period of stagnation of the creation of new UC for the Amazon and Cerrado @fig-uc-timeseries.

```{r}
#| label: fig-uc-timeseries
#| fig-cap: "Time series of Conservation Units."
#| fig-alt: ""
#| fig-asp: 0.70
#| echo: false
#| dev: "ragg_png"

uc_timeseries
```

#### Policies

There are a large number of laws, decrees, resolutions and commitments made in Brazil with the objective to decrease NVS. These are organized in a timeline with some key policies related to NVS @fig-pp-timeline.

```{r}
#| class: scrolling
#| label: fig-pp-timeline
#| fig-cap: "Time line of policies related to Natural Vegetation Suppression."
#| fig-alt: ""
#| fig-asp: 1.8
#| echo: false
#| dev: "ragg_png"

pp_timeline
```

The details of these policies are summarized in the @tbl-timeline, which shows the beggining and end of the policy, their spatial coverage and a general description.

```{r}
#| label: tbl-timeline
#| tbl-cap: Table with policies related to Natural Vegetation Suppression.
#| echo: false
#| dev: "ragg_png"

htmltools::includeHTML("./figs/policies_table.html")

```

## Final results

The final results section will present the divergences and convergences between the qualitative and quantitative results, summarizing the findings in tables and diagrams.

# References {.unnumbered}