# Results

```{r}
#| label: results-load
#| include: false

file_names <- 
  list.files(
    path = "figs", 
    pattern = "rdata", 
    full.names = TRUE,
    recursive = TRUE
  )

lapply(file_names, load, .GlobalEnv)

```

## Theoretical Model

## Sociological Analysis

## Statistical Analysis

### Exploratory Data Analysis

In this section I analyse each variable of my model. 
It is important to note that all quantitative analysis were performed in the data stored in the base grid.
So the figures and tables do not show distributions of the raw observations/estimates, but the values that were transformed to fit the grid cells (@fig-base-grid). 

```{r}
#| label: fig-base-grid
#| fig-cap: Model grid, in which variables were transformed into.
#| fig-asp: 1
#| echo: false

viz_grid_area

```

All visualizations about the variables follows a common color scale, to facilitate their interpretation along the document (@fig-color-scale). The minimum values are represented by light grey color, values close to the median shows a yellow color, and maximum values as dark red color.

```{r}
#| label: fig-color-scale
#| fig-cap: Color scale used to visualize the variables properties.
#| fig-asp: 0.2
#| echo: false

color_scale

```

### Deforestation

A recent research reviewed 65 reports, and found 15 definitions of deforestation, and 9 definitions of forests [@oca-2021].

::: {.callout-important collapse="true"}

READ:

- https://ssrn.com/abstract=4063325
- https://doi.org/10.1086/597794 
- https://doi.org/10.1111/j.1475-682X.2007.00200.x

:::

```{r}
#| label: tbl-deforestation_summary
#| tbl-cap: "Summary table of deforestation."
#| echo: false

htmltools::includeHTML("./figs/eda/deforestation_table.html")

```

```{r}
#| label: fig-deforestation-hist
#| fig-cap: "Distribution of deforestation in the Amazon and Cerrado. The horizontal scale is transformed to base 10 log."
#| fig-asp: 0.70
#| echo: false

deforestation_histogram

```

```{r}
#| label: fig-deforestation-cumsum
#| fig-cap: "Cumulative percent of the deforestation area."
#| fig-asp: 0.70
#| echo: false

deforestation_cumulative

```

```{r}
#| label: fig-deforestation-spatial-distribution
#| fig-cap: "Spatial distribution of the deforestation area."
#| fig-asp: 1
#| echo: false

deforestation_spatial

```

```{r}
#| label: fig-deforestation-timeseries
#| fig-cap: "Time series of deforestation area."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_timeseries

```

```{r}
#| label: fig-deforestation-colsum-natural
#| fig-cap: "Deforested area of different land cover."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_natural

```

```{r}
#| label: fig-deforestation-colsum-human
#| fig-cap: "Area of established LULC after deforestation."
#| fig-alt: "Time series."
#| fig-asp: 0.70
#| echo: false

deforestation_colsum_human

```

## Final results

# References {.unnumbered}